---
title: "Running MACROInFOCUS-like groundwater simulations with macrounchainedFocusGW"
author: "Julien Moeys"
date: "`r Sys.Date()`"
output:
    rmarkdown::html_vignette:
        toc: true
        toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Running MACROInFOCUS-like groundwater simulations with macrounchainedFocusGW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include = FALSE}
verbatim <- function(x){return(paste0("`",x,"`"))} 
anonymise <- function(x){return(macrounchained:::.muc_anonymisePath(x))}
```



Preliminary Note
------------------------------------------------------------

In the text below i use `macrounchained` to indicate 
the [package][r_packages], and `macrounchained()` to indicate 
the function named macrounchained in that package.

More generally, names followed by `()` are function names. 
For example, `macrounchainedFocusGW()` is used for the 
function named macrounchainedFocusGW.

In code examples, calls such as 
`macrounchained::macrounchainedFocusGW()`, i.e. two names 
separated by two colons (`::`), should be understood as 
"a call to the function `macrounchainedFocusGW` in the package 
`macrounchained`".

Users not familiar to R should be aware that it is not possible 
to use backslashes (`\`) as a separator when giving a path. 
Although backslash is the default separator in Windows, 
it is a reserved character in R. Instead, slashes (`/`) 
or double backslashes (`\\`) should be used. As an example, 
defining a path like this: `path <- "C:\swash\macro"` will 
cause an error, while `path <- "C:/swash/macro"` or 
`path <- "C:\\swash\\macro"` are correctly interpreted by R.

In the code examples below, when a path is displayed, 
some Windows variables like `%USERPROFILE%` may appear instead 
of the actual path (something like `C:\Users\username`). This 
is a way to avoid the user name to be displayed in a public 
document. The same is true for the operation log produced by 
`macrounchainedFocusGW()`, as the user name is hidden behind 
relevant Windows variables.



Introduction
============================================================

`macrounchained` is a [package][r_packages] for the program 
[R][r].

`macrounchainedFocusGW()` is a higher level function of the 
package `macrounchained`. It is build on top of the more 
generic function `macrounchained()`. It is specifically 
designed to perform **groundwater** simulations like [MACRO 
In FOCUS][macroinfocus]. 

`macrounchainedFocusGW()` is not as user friendly 
as the graphical user interface (GUI) of MACRO In FOCUS, but 
it allows the simulation of Nth-order metabolites (MACRO In 
FOCUS GUI only allows 1st order metabolites), and of 
large batches of simulations. `macrounchainedFocusGW()` also 
reduces the total number of simulations to be run when simulating 
several degradation products of the same substance, compared 
to MACRO In FOCUS GUI, by avoiding to re-run the same 
"intermediate" simulation multiple time.



Installation and system setup
============================================================

`macrounchained` is based on a collection of four R packages 
([codeinfo][], [macroutils2][], [rmacrolite][] and 
[macrounchained][]). The all need to be installed (see below).

`macrounchained` does not embed any executable for [MACRO][macro] 
or [MACRO In FOCUS][macroinfocus], and neither for the so 
called `exeparfile` tool. So MACRO or MACRO In FOCUS need 
to be installed on the computer for `macrounchained` to work. 
`macrounchained` also needs to know where MACRO or MACRO in FOCUS 
is installed, and how the MACRO-executable and the exeparfile 
are called. `macrounchained` look for this information in 
the following order:

1.  If the path or the executable names have been manually 
    set by the user at the beginning of the R cession (see 
    below), these values are used.
2.  If the path or the executable names are defined as 
    Windows Environment Variables (see below), these values 
    are used.
3.  The default path or the executable names pre-defined 
    in the package are used.

It is therefore possible to use (alternatively) different 
versions of MACRO (for example MACRO In FOCUS or MACRO 5.2) 
on the same computer.



Installation
------------------------------------------------------------

See [the package README page on GitHub](https://github.com/julienmoeys/macrounchained/tree/master/README.md)



Use of Windows Environment Variables to permanently save the path of the directory where MACRO or MACRO In FOCUS are installed
------------------------------------------------------------

It is possible for the user or its system administrator to 
define three Windows Environment Variables that will be used 
by `macrounchained` to know where MACRO or MACRO in FOCUS 
is installed, and how the MACRO-executable and the executables 
are called.

On Windows 10, the Windows Environment Variables can be 
accessed via the Control Panel > System and Security > System > 
Advanced System Settings > Advanced Tab > "Environment Variables..."
System Variables. There, it is possible to define three variables:

*   `rmacrolite_macro_exe`, for example set to 
    `r verbatim(rmacrolite::getRmlPar("macro_exe_default"))`.
*   `rmacrolite_macro_exeparfile`, for example set to 
    `r verbatim(rmacrolite::getRmlPar("macro_exeparfile_default"))`.
*   `rmacrolite_macro_path`, for example set to 
    `r verbatim(anonymise(rmacrolite::getRmlPar("macro_path_default")))`.

Once they have been defined, the values attributed to 
these variables can be accessed by typing:

```{r}
Sys.getenv( "rmacrolite_macro_exe" )
Sys.getenv( "rmacrolite_macro_exeparfile" )
Sys.getenv( "rmacrolite_macro_path" )
```



Alternative to Windows Environment Variables
------------------------------------------------------------

Before using `macrounchainedFocusGW()`, the user can manually 
set the path and/ or the executable names, in R. This setting 
is only valid for the current R cession (i.e. until R is 
closed), and must therefore be re-set for every new R cession.

```{r}
library( "rmacrolite" ) 
```

```{r eval = FALSE}
rmacrolite::rmlPar( "macro_path" = "Macro52Model.exe", 
    "macro_exe" = "exeparfile.exe", 
    "macro_exeparfile" = "C:/swash/macro" ) 
```

You may for example replace `"C:/swash/macro"` with 
`"C:/Program Files (x86)/MACRO52"` if you want to use MACRO 5.2.

As can be seen from the code above, the settings occur to be 
in the package `rmacrolite` which provide all the lower level 
functions used by `macrounchained`.



Running MACRO In FOCUS simulations with macrounchainedFocusGW()
============================================================

Preparing a template par-file
------------------------------------------------------------

A template par-file need to be prepared for each combination 
of:

*   Scenario (ex. Chateaudun)
*   Crop (ex. Winter cereals)

This template need to be prepared with MACRO In FOCUS. Prepare 
a groundwater simulation (with the right scenario and crop), 
and when it is ready, chose the Tools 
menu > Save current run to parameter file and log file. You 
can choose any dummy substance (ex. GW-A) and any dose and 
application date(s) for the template, as long as the DT50 is 
not small.

All the substance parameters, the application date(s) 
and the dose(s), the number of application per application-year, 
and the number of years interval between application-years 
will be overwritten by `macrounchainedFocusGW()`.

In the examples below, I use template par-files that are 
shipped with the packages `macrounchained` or `rmacrolite`



Examples of template par-file used in this document
------------------------------------------------------------

To obtain the path of the template par-files shipped with 
`rmacrolite`, type the following command in R prompt:

```{r results = "hide"}
system.file( package = "rmacrolite", "par-files" )
``` 
```{r echo = FALSE}
anonymise( system.file( package = "rmacrolite", "par-files" ) )
```

and in `macrounchained`

```{r results = "hide"}
system.file( package = "macrounchained", "par-files" )
``` 
```{r echo = FALSE}
anonymise( system.file( package = "macrounchained", "par-files" ) )
```

In the examples below, the path to the par-file is defined 
with a command like

```{r}
pf <- system.file( 
    "par-files", "chat_winCer_GW-X_900gHa_d182.par", 
    package = "rmacrolite" )
```

The result of this command is a character string containing 
the path to the example template par-file called 
`"chat_winCer_GW-X_900gHa_d182.par"`, and save in an R object 
that I called here `pf` (any name is possible). The output 
can be inspected:

```{r results = "hide"}
pf 
``` 
```{r echo = FALSE}
anonymise( pf )
```

The user will most of the time want to use his/ her own 
template, and the code will rather look like:

```{r eval = FALSE}
pf <- "C:/path/to/userparfile.par"
```

where `"C:/path/to/userparfile.par"` is the path and the name 
of a par-file somewhere on the computer.



Preparing a table of substance properties and application patterns
------------------------------------------------------------

About the example: The example given below is for a parent 
substance, called here `GW-M`, a primary metabolite `Met-M1`, 
formed from the degradation of `GW-M`, and a secondary metabolite 
`Met-M2`, formed from the degradation of `Met-M1`.

Notice that it is perfectly possible to run several simulations 
with unrelated substances, i.e. just a batch of simulations 
with different parameters.

In the examples below, the substance is applied every year. 
In this case, the user does not need to indicate the number 
of years interval between application-years, and it will be 
set internally to one year (application every year, the 
default). For applications every other year or every third 
year, an additional column needs to be given in the parameter 
table, `years_interval`, set to 2 or 3, respectively. It can 
vary from substance to substance. An example is given further 
down.



### Options 1: Preparing and importing an Excel file with the properties (with readxl::read_excel)

Some example Excel-files with substance properties and 
application patterns are shipped with `macrounchained`. The 
list of Excel examples can be obtained by typing 
the following command in R prompt:


```{r results = "hide"}
system.file( package = "macrounchained", "xlsx" )
``` 
```{r echo = FALSE}
anonymise( system.file( package = "macrounchained", "xlsx" ) )
```

and the list of files in this folder can be seen with the 
following command:

```{r}
list.files( system.file( package = "macrounchained", "xlsx" ) )
``` 

Copy one of these files and edit the content of sheet `s` (do 
not edit directly the file in the package installation 
directory!).

In order to be able to import an Excel file containing MACRO 
In FOCUS parameters, the package [readxl](https://CRAN.R-project.org/package=readxl) 
needs to be installed. You can check if `readxl` is installed 
by typing the following command in R prompt:

```{r}
(require("readxl"))
``` 

If it returns `TRUE`, the package is installed, and if it 
returns `FALSE`, it is not installed. If needed, it can be 
installed with the following command:

```{r eval = FALSE}
install.packages("readxl")
``` 

The parameters contained in an Excel file can be imported 
with the following command:

```{r}
#   Fetch the path to the Excel file and save it in an R 
#   object called "xl"
xl <- system.file( package = "macrounchained", "xlsx", 
    "macrounchainedFocusGW-with-met.xlsx" )

#   Import the sheet 's' in that Excel file
param1 <- readxl::read_excel( path = xl, sheet = "s" ) 

#   Inspect the content of the table
as.data.frame( param1 ) 
``` 

The reason for `as.data.frame( param1 )` is that `param1` is 
of class `tibble`, which is the format output by 
`readxl::read_excel()`, but not the standard format in R. 
macrounchainedFocusGW() can handle both formats, when it 
comes to input parameters.

A real world Example would rather look like this:

```{r eval = FALSE}
#   Import the sheet 's' in that Excel file
param1 <- readxl::read_excel( 
    path  = "C:/path/to/userexcelfile.xlsx", 
    sheet = "s" )  
``` 

where `"C:/path/to/userexcelfile.xlsx"` is the path and the name 
of an Excel file somewhere on the computer.



### Option 2: Preparing the table of properties in R (with data.frame)

It is of course possible to create the parameter table directly 
in R. Here is how to create the same table as the one imported 
above from an Excel file:

```{r}
param2 <- data.frame( 
    "id"                = c(     1L,       2L,       3L ), 
    "name"              = c( "GW-M", "Met-M1", "Met-M2" ), 
    "kfoc"              = c(     12,     114,        46 ), 
    "nf"                = c(      1,     0.9,       0.9 ), 
    "dt50"              = c(     23,      19,       136 ), 
    "dt50_ref_temp"     = c(     20,      20,        20 ), 
    "dt50_pf"           = c(      2,       2,         2 ), 
    "exp_temp_resp"     = c( 0.0948,  0.0948,    0.0948 ), 
    "exp_moist_resp"    = c(   0.49,    0.49,      0.49 ), 
    "crop_upt_f"        = c(    0.5,     0.5,       0.5 ), 
    "diff_coef"         = c(  5E-10,   5E-10,     5E-10 ), 
    "parent_id"         = c(     NA,      1L,        2L ), 
    "g_per_mol"         = c(    381,     183,       140 ), 
    "transf_f"          = c(     NA,     0.2,       1.0 ), # mol met formed / mol parent degraded
    "g_as_per_ha"       = c(     6,        0,         0 ), 
    "app_j_day"         = c(   309L,    309L,      309L ), 
    stringsAsFactors    = FALSE 
)   
``` 

We can check that the two solutions are identical:

```{r}
all.equal( as.data.frame( param1 ), param2, 
    check.attributes = FALSE ) 
``` 



### Option 3: Preparing the table of properties in R (with read.table)

A 3rd alternative method to prepare a table of parameters 
is to enter the table as a "fixed width format", with a 
variable number of spaces as a column separator. It is not 
the most common way of defining a table in R, but it has the 
advantage of being more human readable.

```{r}
param3 <- read.table( header = TRUE, text = "
    id     name kfoc  nf dt50 dt50_ref_temp dt50_pf exp_temp_resp exp_moist_resp crop_upt_f diff_coef parent_id g_per_mol transf_f g_as_per_ha app_j_day
     1   'GW-M'   12 1.0   23            20       2        0.0948           0.49        0.5     5e-10        NA       381       NA           6       309
     2 'Met-M1'  114 0.9   19            20       2        0.0948           0.49        0.5     5e-10         1       183      0.2           0       309
     3 'Met-M2'   46 0.9  136            20       2        0.0948           0.49        0.5     5e-10         2       140      1.0           0       309
    ", stringsAsFactors = FALSE )
``` 

```{r}
all.equal( as.data.frame( param1 ), param3, 
    check.attributes = FALSE ) 
``` 



Case 1: One application per year, every year
------------------------------------------------------------

In the code below, the path to an example template par-file 
fetched and saved in `pf`, table of properties is 
imported from an example Excel file and saved in `param1`, 
and finally the function `macrounchainedFocusGW`()` is called. 
Notice that this is a dry-run, meaning that all the 
parametrisation is performed, but the three simulations are 
not run. As a consequence, the results of the simulations 
is not analysed either (i.e., the PECs are not calculated).

```{r}
#   Fetch the path to the Excel file and save it in an R 
#   object called "xl"
xl <- system.file( package = "macrounchained", "xlsx", 
    "macrounchainedFocusGW-with-met.xlsx" )

#   Import the sheet 's' in that Excel file
param1 <- readxl::read_excel( path = xl, sheet = "s" ) 

#   Fetch the path to an example template par-file:
pf <- system.file( 
    "par-files", "chat_winCer_GW-X_900gHa_d182.par", 
    package = "rmacrolite" )

library("macrounchained")

#   Dry run: 
res <- macrounchainedFocusGW( 
    s         = param1, 
    parfile   = pf, 
    overwrite = TRUE, 
    run       = FALSE ) # Set to TRUE to run the simulations
``` 

When the simulations are run, that is when the argument 
`run` is set to `TRUE` or omitted (default value is `TRUE`), 
more output will be visible in the operation log, and the 
simulation results is analysed "on the fly" and shown. The 
results will then include something like:

```
<2019-01-23|18:03:43> MACRO runtime 23.13 min
<2019-01-23|18:03:43> Rename MACRO output from MACRO001.BIN to rml_001.bin
<2019-01-23|18:03:43> Analyse simulation results:
$info_general
  conc_percentile rank_period1 rank_period2 method nb_sim_yrs_used
1              80           16           17  focus              20
  nb_yrs_per_period nb_yrs_warmup neg_conc_set_to_0
1                 1             6              TRUE

$info_period
   period_index from_year to_year
1             1      1907    1907
2             2      1908    1908
3             3      1909    1909
4             4      1910    1910
5             5      1911    1911
6             6      1912    1912
7             7      1913    1913
8             8      1914    1914
9             9      1915    1915
10           10      1916    1916
11           11      1917    1917
12           12      1918    1918
13           13      1919    1919
14           14      1920    1920
15           15      1921    1921
16           16      1922    1922
17           17      1923    1923
18           18      1924    1924
19           19      1925    1925
20           20      1926    1926

$water_target_layer_by_period
   period_index     mm_mic mm_mac     mm_tot
1             1 235.470763      0 235.470763
2             2  25.219577      0  25.219577
3             3 249.413842      0 249.413842
4             4 210.212175      0 210.212175
5             5 313.436020      0 313.436020
6             6 203.400054      0 203.400054
7             7 256.735007      0 256.735007
8             8 237.177154      0 237.177154
9             9 142.302469      0 142.302469
10           10 313.514879      0 313.514879
11           11 102.166490      0 102.166490
12           12 133.547061      0 133.547061
13           13 185.370237      0 185.370237
14           14 228.786978      0 228.786978
15           15  45.906557      0  45.906557
16           16   8.512096      0   8.512096
17           17  85.944210      0  85.944210
18           18 113.385342      0 113.385342
19           19  73.593095      0  73.593095
20           20  99.801542      0  99.801542

$solute_target_layer_by_period
   period_index mg_per_m2_mic mg_per_m2_mac mg_per_m2_tot   ug_per_L
1             1  0.0130986457             0  0.0130986457 0.05562748
2             2 -0.0012073002             0 -0.0012073002 0.00000000
3             3  0.0303878063             0  0.0303878063 0.12183689
4             4  0.0326010108             0  0.0326010108 0.15508622
5             5  0.0640291599             0  0.0640291599 0.20428143
6             6  0.0289145618             0  0.0289145618 0.14215612
7             7  0.0360325941             0  0.0360325941 0.14034936
8             8  0.0185321498             0  0.0185321498 0.07813632
9             9  0.0194373138             0  0.0194373138 0.13659154
10           10  0.0301531842             0  0.0301531842 0.09617784
11           11  0.0096938342             0  0.0096938342 0.09488272
12           12  0.0111410579             0  0.0111410579 0.08342421
13           13  0.0120960097             0  0.0120960097 0.06525325
14           14  0.0318550546             0  0.0318550546 0.13923456
15           15  0.0009124727             0  0.0009124727 0.01987674
16           16 -0.0033259016             0 -0.0033259016 0.00000000
17           17  0.0011909045             0  0.0011909045 0.01385672
18           18  0.0020206876             0  0.0020206876 0.01782142
19           19  0.0039686604             0  0.0039686604 0.05392708
20           20  0.0056017143             0  0.0056017143 0.05612853

$water_perc_by_period
   period_index        mm
1             1 236.47974
2             2  19.49554
3             3 250.39844
4             4 200.46625
5             5 318.68542
6             6 202.45740
7             7 256.35547
8             8 237.01562
9             9 153.94165
10           10 307.54321
11           11 107.05884
12           12 127.01587
13           13 186.15771
14           14 234.07593
15           15  38.68237
16           16  15.55029
17           17  88.74512
18           18 109.76221
19           19  71.59131
20           20 102.41870

$conc_target_layer
  ug_per_L ug_per_L_rnd index_period1 index_period2 f_solute_mac f_solute_mic
1 0.139792         0.14            14             7            0            1

--------------------------------------------------------------------------------
<2019-01-23|18:03:43> Simulation 2/5
--------------------------------------------------------------------------------
```

Where `ug_per_L_rnd` in the item `conc_target_layer` is the 
PECgw as MACRO In FOCUS would calculate it (rounded to three 
significant figures).

The commands for a real-world example may rather look like 
this:

```{r eval = FALSE}
param1 <- readxl::read_excel( 
    path  = "C:/path/to/userexcelfile.xlsx", 
    sheet = "s" )

library("macrounchained")

#   Run all the simulations: 
res <- macrounchainedFocusGW( 
    s         = param1, 
    parfile   = "C:/path/to/userparfile.par", 
    overwrite = TRUE ) 
``` 



Case 2: Several application per year, every year
------------------------------------------------------------

Compared to the case above (one application per year, every 
year), the case of several application per year, every year 
will:

*   Need a template par-file that contains the correct 
    number of applications per year (here two).
*   Differ in the way the columns `g_as_per_ha` and 
    `app_j_day` in the parameter table are specified.
    
The example Excel file 
`macrounchainedFocusGW-multiple-applications.xlsx` shows 
how this must be done in Excel (enter several values in the 
same cell by separating with with a vertical bar `|`).

In R, the parameter table can be entered as follow:

```{r}
param2 <- data.frame(
    id             = c(          1,         2 ), 
    name           = c(     "GW-C",   "Met-C" ), 
    kfoc           = c(        172,        52 ), 
    nf             = c(        0.9,       0.9 ), 
    dt50           = c(         20,       100 ), 
    dt50_ref_temp  = c(         20,        20 ), 
    dt50_pf        = c(          2,         2 ), 
    exp_temp_resp  = c(      0.079,     0.079 ), 
    exp_moist_resp = c(       0.49,       0.7 ), 
    crop_upt_f     = c(        0.5,       0.5 ), 
    diff_coef      = c(      5e-10,     5e-10 ), 
    parent_id      = c(         NA,         1 ), 
    g_per_mol      = c(        200,       150 ), 
    transf_f       = c(         NA,   .53/.75 ), 
    g_as_per_ha    = c( "1000|950",     "0|0" ), 
    app_j_day      = c(  "298|305", "298|305" ), 
    stringsAsFactors = FALSE ) 
``` 

The columns `g_as_per_ha` and `app_j_day` now contains 
character strings in which several values have been written, 
separated by a vertical bar (`|`). `macrounchainedFocusGW()` 
will internally convert these into numeric values.

```{r}
param3 <- read.table( header = TRUE, text = "
    id    name kfoc  nf dt50 dt50_ref_temp dt50_pf exp_temp_resp exp_moist_resp crop_upt_f diff_coef parent_id g_per_mol  transf_f     g_as_per_ha app_j_day
     1  'GW-C'  172 0.9   20            20       2         0.079           0.49        0.5     5e-10        NA       200        NA    '1000|950'   '298|305'
     2 'Met-C'   52 0.9  100            20       2         0.079           0.70        0.5     5e-10         1       150 0.7066667         '0|0'   '298|305'
    ", stringsAsFactors = FALSE )
``` 

In the example below, the parameter table is imported from 
Excel and `macrounchainedFocusGW()` is called with these 
parameters and a relevant par-file:

```{r}
#   Fetch the path to the Excel file and save it in an R 
#   object called "xl"
xl <- system.file( package = "macrounchained", "xlsx", 
    "macrounchainedFocusGW-multiple-applications.xlsx" )

#   Import the sheet 's' in that Excel file
param1 <- readxl::read_excel( path = xl, sheet = "s" ) 

#   Fetch the path to an example template par-file:
pf <- system.file( "par-files", 
    "chat_winCer_GW-X_900gHa_d182.par", 
    package = "rmacrolite" )

library("macrounchained")

#   Dry run: 
res <- macrounchainedFocusGW( 
    s         = param1, 
    parfile   = pf, 
    overwrite = TRUE, 
    run       = FALSE ) # Set to TRUE to run the simulations
``` 



Case 3: one application per year, every two years
------------------------------------------------------------

Compared to the first case (one application per year, every 
year), the case of one application per year, every two years 
will only differ by the input par-file. It needs to contain 
the correct number of years without application in between 
years with application.

```{r}
#   Fetch the path to the Excel file and save it in an R 
#   object called "xl"
xl <- system.file( package = "macrounchained", "xlsx", 
    "macrounchainedFocusGW-every-other-year.xlsx" )

#   Import the sheet 's' in that Excel file
param1 <- readxl::read_excel( path = xl, sheet = "s" ) 

#   Fetch the path to an example template par-file:
pf <- system.file( "par-files", 
    "chat_winCer_GW-X_900gHa_d182.par", 
    package = "rmacrolite" )

library("macrounchained")

#   Dry run: 
res <- macrounchainedFocusGW( 
    s         = param1, 
    parfile   = pf, 
    overwrite = TRUE, 
    run       = FALSE ) # Set to TRUE to run the simulations
``` 



System information (for traceability & reproducibility)
============================================================

As visible in the operation log outputs of 
`macrounchainedFocusGW()`, above, the package `macrounchained` 
gives quite extensive information on R-version, the system 
on which the code was run, and not least on the version 
of the different key R packages. Besides the package version, 
are also returned package revision and a package MD5-checksum. 
The version is defined by the package developer(s). The 
revision is machine generated [hash](https://git-scm.com/book/en/v2/Git-Tools-Revision-Selection) 
that can be related to a version of the source code on GitHub. 
The package [MD5 checksum](https://en.wikipedia.org/wiki/MD5) 
can be used to make sure that the original code was not altered 
on the computer used for the calculations.

In `macrounchainedFocusGW()` operation log is also indicated 
`Call from package: 'macrounchained'`. The later would be 
different (likely `'R_GlobalEnv'` instead of `'macrounchained'`) 
if the function `macrounchainedFocusGW()` would have been defined 
by an R script instead of a package.

Finally, the operation log  also contains the version and the 
MD5 checksum of the MACRO executable that was used to run 
the simulations. 

None of these features are fully error- or fool-proof, but 
they altogether should provide a rather robust way to trace 
what version of  `macrounchainedFocusGW()` and MACRO were 
used to generate the simulation results.

The user can obtain the same information by using the function 
`codeinfo()` in the package `codeinfo`:

```{r}
library("codeinfo")

codeinfo( r = TRUE, packages = c( "macrounchained", 
    "rmacrolite", "macroutils2", "codeinfo" ) ) 
``` 



<!-- LIST OF LINKS -->

[r]: https://www.r-project.org/ "The R Project for Statistical Computing"
[r_packages]: https://en.wikipedia.org/wiki/R_(programming_language)#Packages "R packages (Wikipedia)"
[macro]: https://www.slu.se/en/Collaborative-Centres-and-Projects/centre-for-chemical-pesticides-ckb1/models/macro-52/ "MACRO 5.2 (SLU/CKB)"
[macroinfocus]: https://esdac.jrc.ec.europa.eu/projects/macro "MACRO In FOCUS (FOCUS DG SANTE)"
[slu]: https://www.slu.se/ "Swedish University of Agricultural Sciences (SLU)"
[ckb]: https://www.slu.se/ckb "Centre for Chemical Pesticides (CKB)"
[codeinfo]: https://github.com/julienmoeys/codeinfo "R package codeinfo (GitHub)"
[macroutils2]: https://github.com/julienmoeys/macroutils2 "R package macroutils2 (GitHub)"
[rmacrolite]: https://github.com/julienmoeys/rmacrolite "R package rmacrolite (GitHub)"
[macrounchained]: https://github.com/julienmoeys/macrounchained "R package macrounchained (GitHub)"
[macroutils]: https://github.com/julienmoeys/macroutils "R package macroutils (GitHub)"
[github]: https://github.com/ "GitHub development platform"
[focusdgsante]: https://esdac.jrc.ec.europa.eu/projects/focus-dg-sante "FOrum for the Co-ordination of pesticide fate models and their USe"
[devtools]: https://CRAN.R-project.org/package=devtools "R package devtools (CRAN)"
[cran]: https://cran.r-project.org/ "The Comprehensive R Archive Network" 
[kemi]: https://www.kemi.se/en "Swedish Chemicals Agency"
